When are Fuzzy Extractors Possible?

### 2014-10-02 trunk/info-bounds/FuzzyExtPossibility.tex
edits to the intro

\section{Introduction}
\anote{should our results be seen as justifying the distribution specific approach?}
\anote{make the point that narrowing the family $\mathcal{W}$ makes the security guarantees much more brittle}
Sources of secret random bits are necessary for most cryptographic applications.  In many situations, these bits are not explicitly and reliably stored for future use, but are reproduced as a result of some process (such as reading a biometric or a physically unclonable function).
However, many such sources present a problem: noise~\cite{daugman2004,zviran1993comparison,brostoff2000passfaces,ellison2000protecting,mayrhofer2009shake,monrose2002password,pappu2002physical,tuyls2006puf,gassend2002silicon,suh2007physical,bennett1988privacy}.  That is, when a secret is read multiple times, readings are close~(according to some metric) but not identical.  To utilize such sources it is often necessary to remove noise, and achieve the same value in subsequent readings.  \bnote{I tried to remove entropy from this but it needs to be cleaned up}

Bennett, Brassard, and Robert~\cite{bennett1988privacy} introduced this problem in the interactive setting. \lnote{I just realized (after a talk at Allerton) that we are neglecting Wyner's 1975 wiretap channel and follow-up work, which also studied this problem}  They identify two fundamental tasks.  The first, called information-reconciliation, removes the noise without leaking significant information. The second, known as privacy amplification, converts the high entropy secret to a uniform random value.  In this work, we consider the non-interactive version of these problems, in which these tasks are performed with a single message.

The non interactive setting is modeled by a primitive called a fuzzy extractor~\cite{DBLP:journals/siamcomp/DodisORS08}.  %Consider some distribution $W$.%The goal of a fuzzy extractor is to derive stable keys from a distribution $W$ whenever the two readings~(denoted $w$ and $w'$ respectively) are within distance $t$.  
A fuzzy extractor consists of two algorithms: generate~($\gen$) which takes an initial reading $w$ and produces $\key$ along with a nonsecret helper value $p$.  The reproduce~($\rep$) algorithm takes the subsequent reading $w'$ along with the helper value $p$ to reproduce $\key$.   The correctness guarantee is that the key is reproduced if the distance between $w$ and $w'$ is at most $t$.

The security requirement for fuzzy extractors that is $key$ uniform even to an adversary who has observed $p$.   Achieving a strong $\key$ is harder as the allowed error tolerance $t$ increases. It becomes easier for the adversary to guess a $w'$ within distance $t$ of $w$ and an adversary who guesses a nearby $w'$ can obtain $\key$ by running $\rep$.

\anote{Suggested title: "Fuzzy Minentropy"}

\paragraph{A New Entropy Notion}
We introduce a new entropy notion that precisely measures how hard it is for the adversary to guess a value within distance $t$ of the original reading $w$. Suppose $w$ is sampled from a distribution $W$. If the adversary only had access to $\rep$, they would recover $\key$ with maximal probability by guessing $w'$ that maximizes the total probability mass of $W$ within a ball $B_t(w')$ of radius $t$ around $w'$. We  define \emph{fuzzy min-entropy} $\Hfuzz(W) \eqdef -\log \max_w' \Pr[W\in B_t(w')]$.

An adversary who can guess $w'$ can recover the $\key$, that is, $\Hav(\key|p) \le \Hfuzz(W)$.  Super-logarithmic fuzzy min-entropy is necessary for nontrivial key extraction~(\lemref{lem:fuzz necessary}). 
Most known fuzzy extractors obtain security proportional to $\Hoo(W)-\log |B_t(\cdot)|$.  
Observe that, $\Hoo(W)-\log |B_t(\cdot)| \le \Hfuzz(W)$.  However, the difference between these quantities is very large for practical settings.  As an example, iris scans~\cite[Section 5]{blanton2009biometric} appear to have significant $\Hfuzz(W)$ but the value $\Hoo(W) -\log |B_t|<0$.\footnote{Iris scans appear to have $\Hfuzz$ because all iris scans for different people appear to be well-spread in the metric space~\cite{daugman2006probing} and thus $\Hfuzz(W) = \Hoo(W)$.}
We ask: \emph{is fuzzy min-entropy sufficient for fuzzy extraction?}

\anote{Suggested title: "Characterization  for Distribution-aware Extractability". Emphasize if we know the distribution on $W$}

\paragraph{A New Characterization}
We show that for every source $W$ with super-logarithmic $\Hfuzz(W)$, it is possible to construct a fuzzy extractor with a super-logarithmic length $\key$ (\corref{cor:extension to fuzz ext}). We thus show that $\Hfuzz(W)$ is a necessary and sufficient condition for building a fuzzy extractor for a distribution $W$.  While our constructions are not polynomial-time (they are dependent on knowing the complete distribution of $W$), we discuss  how in related settings polynomial-time constructions that provide only computational (rather information-theoretic) security are possible in \secref{sec:related settings}.

>>>>

\section{Introduction}
Sources of reproducible secret random bits are necessary for many cryptographic applications.  In many situations, these bits are not explicitly stored for future use, but are obtained by repeating the same process  (such as reading a biometric or a physically unclonable function) that generated them the first time.
However, bits obtained this way present a problem: noise \cite{daugman2004,zviran1993comparison,brostoff2000passfaces,ellison2000protecting,mayrhofer2009shake,monrose2002password,pappu2002physical,tuyls2006puf,gassend2002silicon,suh2007physical,bennett1988privacy}.  That is, when a secret is read multiple times, readings are close~(according to some metric) but not identical.  To utilize such sources it is often necessary to remove noise, in order to derive the same value in subsequent readings.

The same problem occurs in the interactive setting, in which the secret channel used for transmitting the bits between two users is noisy and/or leaky~\cite{Wyner75}. Bennet, Brassard, and Robert~\cite{bennett1988privacy} identify two fundamental tasks.  The first, called information-reconciliation, removes the noise without leaking significant information. The second, known as privacy amplification, converts the high entropy secret to a uniform random value.  In this work, we consider the non-interactive version of these problems, in which these tasks are performed with a single message.

The non interactive setting is modeled by a primitive called a fuzzy extractor~\cite{DBLP:journals/siamcomp/DodisORS08}.  %Consider some distribution $W$.%The goal of a fuzzy extractor is to derive stable keys from a distribution $W$ whenever the two readings~(denoted $w$ and $w'$ respectively) are within distance $t$.  
A fuzzy extractor consists of two algorithms. The  generate algorithm ($\gen$)  takes an initial reading $w$ and produces $\key$ along with a nonsecret helper value $p$.  The reproduce~($\rep$) algorithm takes the subsequent reading $w'$ along with the helper value $p$ to reproduce $\key$.   The correctness guarantee is that the key is reproduced if the distance between $w$ and $w'$ is at most $t$.

The security requirement for fuzzy extractors that is $\key$ uniform even to an adversary who has observed $p$.   This requirement is  harder to satisfy as the allowed error tolerance $t$ increases, because it becomes easier for the adversary to guess $\key$ by guessing a $w'$ within distance $t$ of $w$ and running $\rep(w',p)$.

\paragraph{Fuzzy Min-entropy}
We introduce a new entropy notion that precisely measures how hard it is for the adversary to guess a value within distance $t$ of the original reading $w$, thus subverting security of $\key$ by running $\rep$. Suppose $w$ is sampled from a distribution $W$.   To have the maximum chance that $w'$ is within distance $t$ of $w$, the adversary would want to maximize the total probability mass of $W$ within a ball $B_t(w')$ of radius $t$ around $w'$.
We  therefore define \emph{fuzzy min-entropy} $\Hfuzz(W) \eqdef -\log \max_w' \Pr[W\in B_t(w')]$.

Super-logarithmic fuzzy min-entropy  is necessary for nontrivial key extraction~(\lemref{lem:fuzz necessary}). 
Most known fuzzy extractors obtain security proportional to $\Hoo(W)-\log |B_t(\cdot)|$.  \lnote{know they don't -- that's better than they achieve, because coding is not that good}
Observe that $\Hoo(W)-\log |B_t(\cdot)| \le \Hfuzz(W) \le \Hoo(W)$.  However, the lower bound is too imprecise (sometimes even negative) for practical settings.  As an example, iris scans~\cite[Section 5]{blanton2009biometric} appear to have significant $\Hfuzz(W)$ but the value $\Hoo(W) -\log |B_t|<0$.\footnote{Iris scans appear to have $\Hfuzz$ because all iris scans for different people appear to be well-spread in the metric space~\cite{daugman2006probing} and thus $\Hfuzz(W) = \Hoo(W)$.}
We ask: \emph{is fuzzy min-entropy sufficient for fuzzy extraction?}

\paragraph{Characterization  for Distribution-aware Extractability}
We show that for every source $W$ with super-logarithmic $\Hfuzz(W)$, it is possible to construct a fuzzy extractor with a super-logarithmic length $\key$ (\corref{cor:extension to fuzz ext}). We thus show that $\Hfuzz(W)$ is a necessary and sufficient condition for building a fuzzy extractor for a distribution $W$.  While our constructions are not polynomial-time (they are dependent on knowing the complete distribution of $W$), we discuss  how in related settings polynomial-time constructions that provide only computational (rather information-theoretic) security are possible in \secref{sec:related settings}.
